---
title: "Reading data"
author: E. Brennan Bollman
date: '`r format(Sys.time(), "%y-%m-%d")`'
output: github_document
---

```{r setup, INCLUDE = FALSE}
library(tidyverse)
library(rvest)
library(httr)

knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = 0.6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom")) 

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d()
scale_fill_discrete = scale_fill_viridis_d()
```


## Scrape a table

I want first table for this page: http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm

First get html, then extract elements of html that I care about

### Read in html.

```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = read_html(url)
```

### Extract the table(s); focus on the first one

```{r}
tabl_marj = 
  drug_use_html %>% 
  html_nodes(css = "table") %>% 
  first() %>% 
  html_table() %>% 
  slice(-1) %>% 
  as_tibble()
```

## Scrape something that's not a table - Star Wars Movie Info

Want starwars movie title, runtime, money. I want the info from [here]: https://www.imdb.com/list/ls070150896/

```{r}
sw_url = "https://www.imdb.com/list/ls070150896/"

sw_html = read_html(sw_url)
```

Grab elements we want. Don't know what CSS tag IMDB uses for movie title.

```{r}
title_vec = 
  sw_html %>% 
  html_nodes(css = ".lister-item-header a") %>% 
  html_text()

gross_rev_vec =
  sw_html %>% 
  html_nodes(css = ".text-muted .ghost~ .text-muted+ span") %>% 
  html_text()

runtime_vec =
  sw_html %>% 
  html_nodes(css = ".runtime") %>% 
  html_text()

swm_df = 
  tibble(
    title = title_vec,
    gross_rev = gross_rev_vec,
    runtime = runtime_vec
  )
```

## API

### NYC Water Data API

For this: https://data.cityofnewyork.us/Environment/Water-Consumption-In-The-New-York-City/ia2d-e54m

```{r}
nyc_water = 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.csv") %>% 
  content("parsed")

nyc_water_json = 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.json") %>% 
  content("text") %>% 
  jsonlite::fromJSON() %>% 
  as_tibble()
```

This gives response on process used, date, status request, and then csv provided. Will need to parse from table in csv to tibble

### CDC BRFSS data

```{r}
brfss_2010 = 
  GET("https://chronicdata.cdc.gov/resource/acme-vg9e.csv") %>% 
  content("parsed")
```

Learn about paging and iterating: https://dev.socrata.com/docs/paging.html

```{r}
brfss_2010_full = 
  GET("https://chronicdata.cdc.gov/resource/acme-vg9e.csv",
      query = list("$limit" = 155000)) %>% 
  content("parsed")
```

## Data that isn't so nice

Pokemon

```{r}
pokemon_data = 
  GET("https://pokeapi.co/api/v2/pokemon/1") %>% 
  content()

pokemon_data$name
pokemon_data$height
pokemon_data$abilities
```

