---
title: "Strings and factors"
author: E. Brennan Bollman
date: '`r format(Sys.time(), "%y-%m-%d")`'
output: github_document
---

```{r setup, INCLUDE = FALSE}
library(tidyverse)
library(rvest)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = 0.6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom")) 

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d()
scale_fill_discrete = scale_fill_viridis_d()
```

## Strings and Regex

Creating string vectors that are not inside a df / tibble. Pattern detection and regular expressions (regex). Detect / replace

```{r}
string_vec_1 = c("my", "name", "is", "jeff")

str_detect(string_vec_1, "jeff")
str_detect(string_vec_1, "e")
str_detect(string_vec_1, "Jeff")

str_replace(string_vec_1, "jeff", "Jeff")
```

Uppercase/lowercase matter here! Can replace with "" to be empty.


```{r}
string_vec_2 = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec_2, "i think")
str_detect(string_vec_2, "^i think")
str_detect(string_vec_2, "i think$")
```

The ^ detects if pattern "starts with". The $ detects if pattern "ends with".

```{r}
string_vec_3 = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec_3, "Bush")
str_detect(string_vec_3, "[Bb]ush")
```

[Bb] can detect either case

```{r}
string_vec_4 = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec_4, "[0-9][a-zA-Z]")
```

Further patterns.

```{r}
string_vec_5 = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec_5, "7.11")
str_detect(string_vec_5, "7\\.11")
```

The "." means will match for any character between the other two

\\. Treats as a dot.

```{r}
string_vec_6 = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec_6, "\\[")
```

\\[ will treat bracket as actual character

stringr package often starts with "str_"

## Factors

```{r}
factor_vec = factor(c("male", "male", "female", "female"))

factor_vec

as.numeric(factor_vec)
```

what happens if I relevel

```{r}
factor_vec = fct_relevel(factor_vec, "male")

factor_vec

as.numeric(factor_vec)
```

All the forcats package starts with "fct_" -- e.g. fct_drop, fct_expand, fct_collapse, fct_relevel, fct_recode (changes label), fct_reorder 

## Strings in dataframe - NSDUH

National survey on drug use and health

```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

tabl_marj = 
  read_html(url) %>% 
  html_nodes(css = "table") %>% 
  first() %>% 
  html_table() %>% 
  slice(-1) %>% 
  as_tibble()
```

Want to remove "a" "be" subscripts. df isn't tidy. 

```{r}
data_marj = 
  tabl_marj %>% 
  select(-contains("P Value")) %>% 
  pivot_longer(
    -State,
    names_to = "age_year",
    values_to = "percent"
  ) %>% 
  separate(age_year, into = c("age", "year"), sep = "\\(") %>% 
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)
  ) %>% 
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West"))) %>% 
  janitor::clean_names()
```

## NSDUH - factors

Factor re-level for a plot.

```{r}
data_marj %>% 
  filter(age == "12-17") %>% 
  mutate(state = fct_relevel(state, "Texas", "Oklahoma")) %>% 
  ggplot(aes(x = state, y = percent, color = year)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Factor re-order from plot.

```{r}
data_marj %>% 
  filter(age == "12-17") %>% 
  mutate(state = fct_reorder(state, percent)) %>% 
  ggplot(aes(x = state, y = percent, color = year)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


